Normalization by Evaluation for Sized Dependent Types

### 2017-02-27 nbe/icfp17.tex

% \cite{boveKraussSozeau:mscs16}
Consequently, each of the aforementioned languages based on Type Theory comes with a termination checker, which needs to reject all non-terminating programs, and should accept sufficiently many terminating programs to allow the user to express her algorithms.  In current termination checkers, programs are required to terminate by structural descent \cite{gimenez:guardeddefinitions}; the structural order may be extended to a lexicographic \cite{abelAltenkirch:jfp02} or size-change termination criterion \cite{jones:sizeChange,wahlstedt:PhD}.  This is not a fundamental limitation, since Type Theory allows many functions to be expressed in a structurally recursive manner, if needed by the help of a well-founded relation \cite{nordstrom:genrec}, inductive domain predicates \cite{boveCapretta:mscs05}, or inductive descriptions of the function graph \cite{bove:entcs09}.  However, the syntactic termination check is very sensitive to reformulations of the program and hostile to abstraction \cite{abel:fics12}.

>>>>

% \cite{boveKraussSozeau:mscs16}
Consequently, each of the aforementioned languages based on Type Theory come with a termination checker, which needs to reject all non-terminating programs, and should accept sufficiently many terminating programs to allow the user to express her algorithms.  In current termination checkers, programs are required to terminate by structural descent \cite{gimenez:guardeddefinitions}; the structural order may be extended to a lexicographic \cite{abelAltenkirch:jfp02} or size-change termination criterion \cite{jones:sizeChange,wahlstedt:PhD}.  This is not a fundamental limitation, since Type Theory allows many functions to be expressed in a structurally recursive manner, if needed by the help of a well-founded relation \cite{nordstrom:genrec}, inductive domain predicates \cite{boveCapretta:mscs05}, or inductive descriptions of the function graph \cite{bove:entcs09}.  However, the syntactic termination check is very sensitive to reformulations of the program and hostile to abstraction \cite{abel:fics12}.
